---
title: "README"
author: Jacques Wainer
date: July, 2023
output: md_document
---

# A Bayesian Bradley-Terry model to compare multiple algorithms on multiple data sets



## Installation

The `bbtcomp` package requires the `cmdstanr` package, which is not on CRAN. So you need to install `cmdstanr` manually, following the steps described in <https://mc-stan.org/cmdstanr/>.



After `cmdstanr` is installed, install `bbtcomp` by issuing 

```
# install.packages("remotes")
remotes::install_github("jwainer/bbtcomp")
```

Installing `bbtcomp` installs the `cmdstanr` R package, but it cannot install or verify CmdStan itself (the compiled Stan backend) -- CmdStan is not an R package, so R's installers have no way to know about it. After installing, run this once to confirm CmdStan is actually set up:

```
check_setup()
# CmdStan found at: /path/to/cmdstan-x.y.z
```

If CmdStan isn't found, `check_setup()` prints (or, with `check_setup(raise_on_error = TRUE)`, stops with) clear setup instructions instead of a cryptic error later. `bbtcomp()` / `mcmcbbt()` also run this check automatically and fail fast with the same message if CmdStan is missing.

```{r x0, echo=F, include=F}
library(bbtcomp)
```

## Data format

In general terms, data is organized as a data frame where the rows are
the data sets and the columns are the algorithms. The column names
will be used as the names of the algorithms being compared.

There may be a column that does not contain the results of an algorithm in the data sets, but contain an id for each data set. This is called the `dbcol`.  


In more detail, the data may be:

- a data frame or an array, with no `dbcol`. The column names are the names of the algorithms, and the entry in line $i$ and column $j$ is the measure of some metric of algorithm $j$ on the data set $i$. Usually, the measure is a mean of different evaluations on different test sets. 

- a data frame with a `dbcol` - a column with a string value that identifies the data set name. 

- two matrices, where the column names are the same in both, the first matrix containing the mean (across the cross-validation folds) of the measure for each algorithm and data set, and the second matrix, the standard deviation of the measures (across the same folds). In this case the algorithm can compute the non-paired version of the local ROPE.


The algorithm assumes that higher values of the metric are better, as with accuracy, AUC, F1, and other metrics. But if the metric is an error measure, such as RMSE, MAE, or others, the user should multiply these values by -1 (so that higher values -- negative but close to 0 -- are better). If the algorithm did not run for that fold, or for that data set, the entry should be `NA`.

The `bbtcomp` package comes with a data set comparing 16 algorithms on 132 data sets (as described in the paper). The data set is called `ll`

```{r x1, echo=T}
ll[1:8,]
```

Let us use a smaller sub-data set

```{r x2, echo=T}
ss <- ll[1:80, 1:6] # first 20 data sets (4 folds each) and 5 algorithms
ss
```

This data frame corresponds to the second type of data for the bbtcomp algorithm: a data frame with a column that indicates the data set, and potentially multiple measures for each data set (for multiple folds). 

A data frame that corresponds to the first data format is obtained by:

```{r x2a, echo=T}
library(dplyr)
ssmean <- ll %>% group_by(db) %>% summarize(across(everything(),.fns = mean))
ssmean <- ssmean[,-1]
ssmean
```

This data has only a single measure per algorithm (column) and per data set (each row, implicitly). This format does not allow one to use the local ROPE (discussed below).




Finally, the third format is a pair of data frames, one with the mean measure for each data set and algorithm (the `ssmean` data frame above), and one with the standard deviation of the measures. The second data frame can be computed by:

```{r x2b, echo=T}
sssd <- ll %>% group_by(db) %>% summarize(across(everything(),.fns = sd))
sssd <- sssd[,-1]
sssd
```


## Before you start using bbtcomp

### Running chains in parallel

First, to run the chains in parallel:

```{r y0, echo=T}
options(mc.cores = parallel::detectCores(logical = FALSE))
```

If, when running bbtcomp, you see the message 

```
Running MCMC with 4 sequential chains...
```
(the key word there is **sequential**), it is because you did not set up Stan to run the chains in parallel with the `options` command above.

### Auxiliary folder


CmdStan, which serves as the interface between bbtcomp and Stan (the MCMC engine under the hood of bbtcomp), uses files as an intermediary between the two processes. First, the Bayesian model of BBT is translated into C++ and then compiled and saved into a file. The sampling process then saves the results as .csv files (one per chain). All these files are stored in an auxiliary folder.

This auxiliary folder can be set using 

```{r y00, echo=T}
options(bbtcomp.dir = "~/.bbtcomp")
```

In this case, the auxiliary folder is "~/.bbtcomp". If the folder is not set, then bbtcomp will use a temporary folder that will be deleted as soon as the R session is terminated.

The advantage of setting the auxiliary folder with the `options` command is that the compiled Stan program is stored there and there is no need to compile it again in another R session using bbtcomp. The disadvantage is that each call to `bbtcomp` generates 4 .csv files, and the user should erase these .csv files from time to time.

The advantages and disadvantages of using a temporary folder are the opposite of those of the fixed auxiliary folder: a new compilation of the Stan model will be necessary each time `bbtcomp` is called in a new R session, but there is no need to delete the .csv files from time to time.


## Basic use

Basic bbtcomp operation:


```{r y1, echo=T}
x <- bbtcomp(ss)
```
will return a `bbt model` which is a list with 3 components that contains:

- a `cmdstanr` fit model (with the samples)

- a wintable

- a Boolean stating whether the Davidson model was used 


Let us plot the results:

```{r y2, echo=T}
plot_pwin(x)
``` 

Let us see the results as a table:

```{r y3, echo=T}
table_pwin(x)
``` 

`table_pwin()` returns a data.frame (S3 class `"bbt_pwin_table"`) where the
two algorithm names of each pair are stored as separate `larger`/`smaller`
columns (the printed `pair` column above is just those two combined for
display, e.g. `"lgbm > rf"` means `larger = "lgbm"`, `smaller = "rf"`), so
code can use them directly instead of parsing the printed string back
apart, e.g. `subset(table_pwin(x), larger == "lgbm")`.

Or only the comparisons of the `rf` algorithms with the others
```{r y4, echo=T}
table_pwin(x,control='rf')
``` 

Or other summary of the probability distributions (median, low and high limits of the 95% HDI):

```{r y5, echo=T}
table_pwin(x, columns = c("median", "low", "high"), hdi=0.95)
``` 

For the `ssmean` data format (without the dbcol), use 
```{r y1b, echo=T}
x2 <- bbtcomp(ssmean, dbcol=0)
```

For the `ssmean` and `ssds`  data format (without the dbcol), use 
```{r y1c, echo=T}
x3 <- bbtcomp(ssmean, sssd)
```



### Local ROPE

Although the ss data has 4 entries for each data set -- the results on 4 different test sets (4-fold cross-validation) -- the results displayed above compute the mean for each algorithm and data set and perform the BBT on the mean results. 

The paper discusses that using the fold data one can convert some of the victories of one algorithm over another into a tie, because the difference of the means is smaller or much smaller than the variance of the results for the folds. The paper calls it a local ROPE, and discusses two different approaches to the local ROPE, when the measures for the folds are paired (the same test set was measured for all algorithms) or not paired (each test set was potentially different for each algorithm).

To use the local ROPE both the paired and not-paired versions:
```{r y7, echo=T}
y <- bbtcomp(ss,lrope=T, paired=T) # paired version - default 
z <- bbtcomp(ss,lrope=T, paired=F) # not paired version 
``` 

In the case of the ss data in particular, using either version of the local ROPE will not change the number of victories and losses into ties except for one pair of algorithms.

For the other data formats: only the mean data `ssmean` does not allow the computation of the local ROPE. For the mean and standard deviation data (`ssmean` and `ssds`) only the non-paired version of local ROPE can be computed. 


### Convergence check

Convergence check of the sampling is performed by the `convergence_check` function, which only calls the `cmdstan_diagnose` function from CmdStan.

`cmdstanr` uses a directory to store two important pieces of information from the sampling process. The first is the compiled Stan model: the first time one calls `bbtcomp`, the Stan model is compiled and stored in this directory. The directory is set by the `dir` parameter of the `bbtcomp` function, and the default is `.bbtcomp` in the current directory.

The second piece of data stored in the directory is the .csv files containing the samples themselves. By default, 4 chains are sampled in parallel, so the sampling process generates 4 .csv files with the results. The samples themselves are part of the bbt model object returned by `bbtcomp`, and this object is passed to the functions that generate the tables and the graphs. By default, `bbtcomp` deletes the .csv files in the directory since they tend to grow large (each call to `bbtcomp` generates 4 new large .csv files). But the convergence check performed by the `cmdstan_diagnose` function reads those files, so to run the convergence tests one must disable the cleanup of the .csv files.

```{r y8, echo=T}
y <- bbtcomp(ss) 
convergence_check(y) 
``` 


### Posterior Predictive Check

The posterior predictive checks of the model in relation to the data can be generated by

```{r y9, echo=T}
plot_ppc(y) 
``` 

The table version of the same check is obtained as 

```{r y10, echo=T}
table_ppc(y) 
``` 

### The win/loss table

The table of wins and losses (a wintable) for all algorithms can be accessed as the `wintable` component of the model returned by `bbtcomp`. To print the table of victories and losses with the ties shown explicitly (i.e., unprocessed), use:

```{r y11, echo=T}
table_wintable(y$wintable, which = "pre") 
``` 

To print the table with the ties processed, use

```{r y12, echo=T}
table_wintable(y$wintable) 
``` 

To print both tables side by side:

```{r y13, echo=T}
table_wintable(y$wintable, which = "both") 
``` 

## Python implementation

### Installation

The Python implementation needs the `cmdstanpy` package, so please install it following the steps described in <https://mc-stan.org/cmdstanpy/>.

The Python package lives in the `bbtcomp_python` directory as an installable pip package. Install it (in editable mode, while developing) with:

```
pip install -e bbtcomp_python
```

### Using the Python implementation

The Python package mirrors the full R package: `bbtcomp`, `make_wintable`, `mcmcbbt`, `table_pwin`, `table_ppc`, `table_wintable`, `convergence_check`, `get_waic`, `get_loo`, `plot_pwin`, `plot_ppc`, and `load_ll()` for the bundled `ll` data set.

```{python ww2, echo=T, eval=F, python.reticulate = FALSE}
from bbtcomp import bbtcomp, table_pwin, load_ll

ll = load_ll()
``` 

`bbtcomp()` takes the same parameters, data formats, and so on, as the R version. A few differences:

- to indicate that there is no `dbcol`, use `dbcol=None` (or `dbcol=-1`);

- to control the auxiliary folder, set the `output_dir` parameter. If left unset, `bbtcomp` uses a temporary folder that is deleted as soon as the Python session ends.

See `bbtcomp_python/README.md` for full usage examples.

